---
title: "HW4"
author: "Genavieve Middaugh"
date: "2025-02-14"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r global, echo= FALSE}
library(tidyverse)
library(mosaic)


```

## **Problem 1 - Iron Bank**

The Securities and Exchange Commission (SEC) is investigating the Iron Bank, where a cluster of employees
have recently been identified in various suspicious patterns of securities trading that violate federal “insider
trading” laws. Here are few basic facts about the situation:
• Of the last 2021 trades by Iron Bank employees, 70 were flagged by the SEC’s detection algorithm.
• But trades can be flagged every now and again even when no illegal market activity has taken place.
In fact, the SEC estimates that the baseline probability that any legal trade will be flagged by their
algorithm is 2.4%.
• For that reason, the SEC often monitors individual and institutional trading but does not investigate
incidents that look plausibly consistent with random variability in trading patterns. In other words,
they won’t investigate unless it seems clear that a cluster of trades is being flagged at a rate significantly
higher than the baseline rate of 2.4%.
Are the observed data (70 flagged trades out of 2021) consistent with the SEC’s null hypothesis that, over
the long run, securities trades from the Iron Bank are flagged at the same 2.4% baseline rate as that of other
traders?
Use Monte Carlo simulation (with at least 100000 simulations) to calculate a p-value under this null hypothesis.
Include the following items in your write-up:
• the null hypothesis that your are testing;
• the test statistic you used to measure evidence against the null hypothesis;
• a plot of the probability distribution of the test statistic, assuming that the null hypothesis is true;
• the p-value itself;
• and a one-sentence conclusion about the extent to which you think the null hypothesis looks plausible
in light of the data. This one is open to interpretation! Make sure to defend your conclusion

## **Problem 2 - Health Inspections**

The local Health Department is investigating a popular local restaurant chain, Gourmet Bites, after receiving
a higher-than-usual number of health code violation reports. Here are a few key points about the situation:
• Over the last year, 1500 health inspections were conducted across various restaurants in the city, with
various branches of Gourmet Bites inspected a total of 50 times.
• Of these 50 inspections, 8 resulted in health code violations being reported.
• Typically, the Health Department’s data shows that, on average, 3% of all restaurant inspections result
in health code violations due to random issues that can occur even in well-managed establishments.
The Health Department wants to ensure that any action taken is based on solid evidence that Gourmet Bites’
rate of health code violations is significantly higher than the citywide average of 3%.
Question: Are the observed data for Gourmet Bites consistent with the Health Department’s null hypothesis
that, on average, restaurants in the city are cited for health code violations at the same 3% baseline rate?
Use a Monte Carlo simulation (with at least 100,000 simulations) to calculate a p-value under this null
hypothesis. Follow the same answer format as in the prior problem.


## **Problem 3 - Evaluating Jury Selection for Bias**



